\thispagestyle{empty} 
\SpecialTitle{Abstract}
\noindent Probabilistic forecasting is vital in time series analysis, with not just point forecasts but quantification of uncertainty in the shape of prediction intervals in applications like finance, energy, web traffic, etc., where decision-making under risk is crucial. Methods like Traditional and Advanced Lower Upper Bound Estimation (LUBE) are sensitive to loss functions and prone to overfitting in intricate temporal patterns. Quantile Regression (QR-based) and Bootstrap-based methods are computationally intensive. Parametric methods, such as Gaussian and Weibull distributions, have restrictive structural shapes that can become incapable of accounting for the underlying uncertainty in the residuals. The limitation of individual probabilistic methods can be overcome by developing hybrid methods, integrating the merits of multiple methods. Motivated by this, two hybrid methods namely, LUBE-Weibull and LUBE-QR are proposed employing deep learning models for probabilistic time series forecasting. These methods overcome the limitations of traditional methods by incorporating the strengths of LUBE, Advanced LUBE, Weibull distribution, and QR-based methods. The first hybrid method takes the loss function of the LUBE method and adds to it a Weibull distribution-based residual correction, while the second hybrid method takes the loss function of the Advanced LUBE method and adds to it a QR-based residual correction, thus achieving very good results. LUBE-Weibull based hybrid method achieved 100\% PICP values with reduced PINAW values while being computationally efficient while LUBE-QR based method consistently achieved ideal PICP values across five datasets and confidence levels while having the lowest PINAW and ACE values. It is also statistically determined by Friedman Nemenyi hypothesis test to be a better method than the traditional ones and the LUBE-Weibull based hybrid method.

\vspace{5mm}
\noindent\textit\textbf{{Keywords}}:~\textit{Probabilistic Forecasting};~\textit{Prediction Intervals};~\textit{Advanced LUBE Method};~\textit{Quantile Regression (QR)-Based Method};~\textit{Deep Learning Models (LSTM, CNN, GRU, BiLSTM)};~\textit{Interval Evaluation Metrics (PICP, PINAW, AWE, ACE)};~\textit{Non-Parametric Methods}.